Leadership across cultures and contexts: an exploratory analysis

### 2020-06-18 leadership_across_cultures_contexts.Rmd
Updated TLQ special issue citations.

<!-- EVOLUTIONARY THEORY APPROACH -->
<!-- Substantial scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover interrelationships of dimensions of leadership, including the traits of leaders, their roles, and relationships with culture, economics, and group contex. Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology. Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2019; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @alami_mothers_2020; @cheng_two_2013; @wiessner_collective_2019; @redhead_dynamics_2018; @garfield_investigating_2019; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2018; @lonati_what_2019]. This integration as been informative [see @van_vugt_genes_2020]. -->

@@

## The functions of leadership across cultures and contexts
<!-- INTRO -->
@roscoe_social_2009 proposed leadership in small-scale societies is structured by the joint demands of facilitating within-group conflict resolution and cooperation, and signaling coalitional strength to rival groups. Among many small-scale societies, which generally lack institutionalized political structures, leadership at the community level is typically informal and leaders emerge in response to group demands [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993]. Theoretical literature suggests these demands most commonly arise in the contexts of collective actions, resolving conflicts between families, and punishing norm violations [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2019]. Leaders also play important roles in ritual contexts, such as in initiations [@singh_cultural_2017], facilitating marriages [@walker_evolutionary_2011], and organizing feasts [@wiessner_food_1996]. Between-group interactions, including both conflict and cooperation, are also common among hunter-gatherers and other nonindustrial societies [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012].

@@

<!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
Intelligent and knowledgeable individuals [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], elders, and respected individuals tend to have greater influence over community decisions across cultures, including among egalitarian hunter-gatherers, stratified chiefdoms, and nation state organizations [@silverman_how_1978; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990; @garfield_evolutionary_2019; @garfield_investigating_2019]. Leaders also tend to have large social networks [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012], and embody the qualities most respected by the group [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019]. Leaders across multiple contexts tend to be taller and more physically formidable, which might facilitate resolving inter- and intra-group  conflicts [@lukaszewski_role_2016; @von_rueden_leadership_2014; @von_rueden_roots_2014]. Charisma and oratory skills are common properties of leaders when they must rely on persuasion to influence community members [@grabo_charismatic_2016].

@@

<!-- MALE BIAS -->
Substantial evidence suggests a near universal male-bias in community leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2018; @tiger_female_1978]. Sex, however, is confounded with factors such as social network size, economic specialization, and education, that are also associated with leadership. After controlling for these factors, some studies have found that sex is not a strong predictor of community leadership [@yanca_female_2004; @garfield_investigating_2019; @von_rueden_sex_2018].

@@

<!-- Leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019]. Much of human leadership goes beyond group coordination problems, such as leaders important roles in devising strategic outcomes between individuals, offering advice or decision-making services to individuals, and in making direct prosocial investments that do not involve substantial coordination. We therefore suggest leadership dynamics emerge from inter-individual competencies and leaders' abilities to provide benefits *or* impose costs, with these dimensions interacting with market forces rooted in follower preferences, group context, and cultural norms.  -->

 <!-- Whereas @pietraszewski_evolution_2019 emphasizes leaders ability to propose various enterprises and followers ability to evaluate them, -->

@@

These potentially universal dimensions of leadership correspond to those highlighted by several evolutionary theories that have drawn heavily on select ethnographic cases [@garfield_review_2019]. They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], for example, which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigious leaders. We labeled one of our leader quality features "cultural conformity", however, in which leaders are culturally conservative and exemplify desirable traits such as honesty, humility, and "proper behavior" (Figure \@ref(fig:cluster)). This feature might correspond, at least in part, to leaders who serve as models for their followers [see also @garfield_investigating_2019].

@@

When ethnographic texts described female leaders exclusively, they were relatively more likely to be described as married to a high-quality spouse in a polygynous marriage, and the recipients of various mating benefits (Figure S\@ref(fig:heatmapsex)). Female leaders also had higher levels of evidence for the *Social, reproductive, and material success* "feature" variable (Figure S\@ref(fig:heatmapfeaturesex)). Evidence for female leaders was also more likely to be found in kin groups and residential subgroups. This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985; @yanca_female_2004]. @garfield_investigating_2019 similarly found evidence of positive assortative mating of high status leaders in a small-scale society with relative gender-egalitarianism. Our results are also consistent with theoretical arguments emphasizing women as leaders of families [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019].

@@

In our view, social context is critical to these discussions [see @wiessner_power_2010]. Leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social "statuses." In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who manage the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012]. @garfield_evolutionary_2019 found that leader organization dimensions clustered separately from leader prestige dimensions, suggesting they are distinct. @garfield_investigating_2019 found that some leaders among recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect. Some older individuals, on the other hand, were highly respected but had little social influence or authority.

@@

@henrich_evolution_2001 introduced this important distinction to social science disciplines outside of anthropology, agreeing with @tiger_dominance_1970 and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] on a primate heritage of dominance, but offering an evolutionary account of prestige rooted in social rather than sexual selection for culturally acquired competencies. @henrich_evolution_2001 inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2019].

@@

Leaders, possibly due to their special competencies, are able to devise and implement strategic solutions to group problems including resolving conflicts within the group (the most widely supported function in our sample) and providing counsel (also widely supported), which again are functions that are commonly noted in the anthropological and psychological leadership literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002]. The feature of leader organizational functions aligns with many evolutionary theories that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2019; for review, see @garfield_evolutionary_2019].

@@

Our revised model proposes that leaders deploy cognitive, social, material, and somatic capital to provide benefits and impose costs on followers. If their abilities to provide benefits or impose costs are similar to other groups members, as in "chiefs-for-a-day", they acquire little status, but if their abilities are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2019]. There was particularly strong evidence for cognitive and social capital, with cognitive capital associated with both providing benefits and imposing costs, particularly among shaman leaders (Figure \@ref(fig:fivefold)).

@@

<!-- DIVERSE FUNCTIONS -->
<!-- Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2019; @price_evolution_2014]. Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group. However, the ubiquity of conflict resolution by leaders across social contexts and cultures calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits [@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018; @hagen_leadership_2019]. -->

@@

<!-- SUMMARY AND SUGGEST LEADERSHIP BE PART OF OTHER GROUP RESEARCH -->
<!-- These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and levels of social organization. Political leadership is common across a broad range of populations, including among relatively small-scale and egalitarian hunter-gatherers. Among hunter-gatherers, however, leaders more frequently emerge within residential subgroups and kin groups. Given the vast majority of human evolutionary history involved subsistence economies characterized by hunting and gathering and greater residential mobility [@lee_cambridge_1999; @marlowe_hunter-gatherers_2005], these results suggest leadership within kin groups and local subgroups likely has a long evolutionary history. Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2018; @von_rueden_sex_2018].  -->

@@

**WEIRD leadership costs and benefits**. Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2019], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation. In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs. @rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in WEIRD urban neighborhood organizations, which can be viewed as self-interested groups that face group-wide problems and necessarily share public benefits. The costs associated with leadership in neighborhood organizations follow those associated with other collective enterprises (e.g., investments of time and energy in management and maintenance of group coordination). @rich_dynamics_1980 conceptualizes four categories of the benefits associated with neighborhood organization. Leaders of neighborhood organizations (1) may receive benefits of the public goods produces from collective actions (as other non-leaders would as well); (2), may receive selective benefits from surpluses produced through collective actions; (3) may receive incidental benefits from their leadership role; and (4) leaders of neighborhood organizations may enjoy their service in the position. @rich_dynamics_1980 interviewed 54 officers from 11 metropolitan neighborhood organizations in the Midwest US. Results from extensive interviews suggest personal fulfillment and enjoyment are the main motivational drivers of participation in leadership roles of neighborhood organizations, suggesting the process of leadership for individuals is distinct from theoretical models rooted in economics. Therefore, a primary benefit of leadership, in some contexts, is likely to be intrinsic personal fulfillment and the opportunity to execute personal values of community service.

>>>>

<!-- EVOLUTIONARY THEORY APPROACH -->
<!-- Substantial scientific work across psychology, sociology, political science, and managerial studies has focused on trying to uncover interrelationships of dimensions of leadership, including the traits of leaders, their roles, and relationships with culture, economics, and group contex. Until recently, this large body of work was relatively disconnected from approaches in anthropology and biology. Under a unifying evolutionary framework, social and biological scientists have now produced a sizable body of theory on the evolution and cross-cultural patterning of leadership and followership [e.g., @price_evolution_2014; @henrich_big_2015; @richerson_tribal_2009; @van_vugt_evolutionary_2006; @pietraszewski_evolution_2020; @hooper_theory_2010; @kaplan_evolutionary_2009; @von_rueden_roots_2014; @hagen_leadership_2019], accompanied by a growing body of empirical results [e.g., @bowser_womens_2010; @alami_mothers_2020; @cheng_two_2013; @wiessner_collective_2019; @redhead_dynamics_2018; @garfield_investigating_2020; @garfield_evolutionary_2019; @von_rueden_leadership_2014; @von_rueden_sex_2018; @smith_leadership_2016; @smith_obstacles_2020; @lonati_what_2019]. This integration as been informative [see @van_vugt_genes_2020]. -->

@@

## The functions of leadership across cultures and contexts
<!-- INTRO -->
@roscoe_social_2009 proposed leadership in small-scale societies is structured by the joint demands of facilitating within-group conflict resolution and cooperation, and signaling coalitional strength to rival groups. Among many small-scale societies, which generally lack institutionalized political structures, leadership at the community level is typically informal and leaders emerge in response to group demands [@von_rueden_leadership_2015; @kantner_identifying_2010; @lewis_leaders_1974; @boehm_egalitarian_1993]. Theoretical literature suggests these demands most commonly arise in the contexts of collective actions, resolving conflicts between families, and punishing norm violations [e.g., @glowacki_leadership_2015; @gavrilets_convergence_2016; @hooper_theory_2010; @von_rueden_solving_2015; @henrich_big_2015; @pietraszewski_evolution_2020]. Leaders also play important roles in ritual contexts, such as in initiations [@singh_cultural_2017], facilitating marriages [@walker_evolutionary_2011], and organizing feasts [@wiessner_food_1996]. Between-group interactions, including both conflict and cooperation, are also common among hunter-gatherers and other nonindustrial societies [@glowacki_evolutionary_2017; @hames_pacifying_2019; @richerson_cultural_2016; @henrich_evolution_2009; @apicella_social_2012].

@@

<!-- MORE UNIVERSAL QUALITIES FROM WEIRD AND NON WEIRD -->
Intelligent and knowledgeable individuals [e.g., @johnson_leadership_2015; @judge_intelligence_2004; @neel_being_1980; @antonakis_does_2009], elders, and respected individuals tend to have greater influence over community decisions across cultures, including among egalitarian hunter-gatherers, stratified chiefdoms, and nation state organizations [@silverman_how_1978; @tooby_cognitive_2006; @garfield_review_2019; @bass_handbook_1990; @garfield_evolutionary_2019; @garfield_investigating_2020]. Leaders also tend to have large social networks [@von_rueden_leadership_2014; @von_rueden_roots_2014; @walker_living_2012], and embody the qualities most respected by the group [@henrich_evolution_2001; @van_vugt_evolutionary_2006; @collier_politics_1981; @low_men_1992; @roscoe_social_2019]. Leaders across multiple contexts tend to be taller and more physically formidable, which might facilitate resolving inter- and intra-group  conflicts [@lukaszewski_role_2016; @von_rueden_leadership_2014; @von_rueden_roots_2014]. Charisma and oratory skills are common properties of leaders when they must rely on persuasion to influence community members [@grabo_charismatic_2016].

@@

<!-- MALE BIAS -->
Substantial evidence suggests a near universal male-bias in community leadership across human societies [@low_sex_1992; @whyte_status_1978; @sanday_female_1981], as well as across diverse nonhuman social species [@smith_obstacles_2020; @tiger_female_1978]. Sex, however, is confounded with factors such as social network size, economic specialization, and education, that are also associated with leadership. After controlling for these factors, some studies have found that sex is not a strong predictor of community leadership [@yanca_female_2004; @garfield_investigating_2020; @von_rueden_sex_2018].

@@

<!-- Leadership and followership stem from biological market forces and information-processing roles [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2020]. Much of human leadership goes beyond group coordination problems, such as leaders important roles in devising strategic outcomes between individuals, offering advice or decision-making services to individuals, and in making direct prosocial investments that do not involve substantial coordination. We therefore suggest leadership dynamics emerge from inter-individual competencies and leaders' abilities to provide benefits *or* impose costs, with these dimensions interacting with market forces rooted in follower preferences, group context, and cultural norms.  -->

 <!-- Whereas @pietraszewski_evolution_2020 emphasizes leaders ability to propose various enterprises and followers ability to evaluate them, -->

@@

These potentially universal dimensions of leadership correspond to those highlighted by several evolutionary theories that have drawn heavily on select ethnographic cases [@garfield_review_2019]. They partially correspond to the prestige model [@henrich_big_2015; @henrich_evolution_2001], for example, which emphasizes knowledge, skill and generosity, but they do not include a key feature of that model: emulation of prestigious leaders. We labeled one of our leader quality features "cultural conformity", however, in which leaders are culturally conservative and exemplify desirable traits such as honesty, humility, and "proper behavior" (Figure \@ref(fig:cluster)). This feature might correspond, at least in part, to leaders who serve as models for their followers [see also @garfield_investigating_2020].

@@

When ethnographic texts described female leaders exclusively, they were relatively more likely to be described as married to a high-quality spouse in a polygynous marriage, and the recipients of various mating benefits (Figure S\@ref(fig:heatmapsex)). Female leaders also had higher levels of evidence for the *Social, reproductive, and material success* "feature" variable (Figure S\@ref(fig:heatmapfeaturesex)). Evidence for female leaders was also more likely to be found in kin groups and residential subgroups. This suite of characteristics is consistent with discussions of high status first-wives who gain and maintain increased social influence across the lifespan by leveraging their extended kin and social networks [@goodale_tiwi_1971; @brown_her_1985; @yanca_female_2004]. @garfield_investigating_2020 similarly found evidence of positive assortative mating of high status leaders in a small-scale society with relative gender-egalitarianism. Our results are also consistent with theoretical arguments emphasizing women as leaders of families [e.g., @garfield_evolutionary_2019; @hagen_leadership_2019].

@@

In our view, social context is critical to these discussions [see @wiessner_power_2010]. Leadership and followership emerge across levels of social organization and individuals simultaneously maintain multiple social "statuses." In many cultures, for example, communities form work-groups and appoint "chiefs-for-a-day" who manage the work but have little status or authority otherwise [e.g., @macfarlan_competitive_2012]. @garfield_evolutionary_2019 found that leader organization dimensions clustered separately from leader prestige dimensions, suggesting they are distinct. @garfield_investigating_2020 found that some leaders among recently settled hunter-gatherers had organizing responsibilities but unexceptional levels of respect. Some older individuals, on the other hand, were highly respected but had little social influence or authority.

@@

@henrich_evolution_2001 introduced this important distinction to social science disciplines outside of anthropology, agreeing with @tiger_dominance_1970 and Barkow [-@barkow_prestige_1975; -@barkow_prestige_1980] on a primate heritage of dominance, but offering an evolutionary account of prestige rooted in social rather than sexual selection for culturally acquired competencies. @henrich_evolution_2001 inspired much subsequent work across the behavioral sciences [e.g., @van_vugt_dual_2019; @cheng_two_2013; @maner_2016; @halevy_status_2012; @suessenbach_dominance_2019; @maner_dominance_2017; @garfield_evolutionary_2019; @garfield_investigating_2020].

@@

Leaders, possibly due to their special competencies, are able to devise and implement strategic solutions to group problems including resolving conflicts within the group (the most widely supported function in our sample) and providing counsel (also widely supported), which again are functions that are commonly noted in the anthropological and psychological leadership literature [@wiessner_risk_1982; @wiessner_collective_2019; @boehm_hierarchy_1999; @knauft_violence_1991; @de_cremer_intergroup_2002]. The feature of leader organizational functions aligns with many evolutionary theories that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; @pietraszewski_evolution_2020; for review, see @garfield_evolutionary_2019].

@@

Our revised model proposes that leaders deploy cognitive, social, material, and somatic capital to provide benefits and impose costs on followers. If their abilities to provide benefits or impose costs are similar to other groups members, as in "chiefs-for-a-day", they acquire little status, but if their abilities are exceptional relative to other group members, they gain status via biological market mechanisms [@hammerstein_biological_2016; @garfield_evolutionary_2019; @hagen_leadership_2019; @pietraszewski_evolution_2020]. There was particularly strong evidence for cognitive and social capital, with cognitive capital associated with both providing benefits and imposing costs, particularly among shaman leaders (Figure \@ref(fig:fivefold)).

@@

<!-- DIVERSE FUNCTIONS -->
<!-- Contemporary theoretical models of the evolution of leadership often emphasize leaders roles in organizing collective action and in managing between group conflicts [@hooper_theory_2010; @pietraszewski_evolution_2020; @price_evolution_2014]. Results here do support the important functions of leaders in organizing cooperation and facilitating social functions -- services which provide direct benefits to the group. However, the ubiquity of conflict resolution by leaders across social contexts and cultures calls for greater attention to the roles of leaders of all types in managing intra-group conflict between group members -- services tailored to provide direct benefits to a few individuals and more indirect group benefits [@glowacki_leadership_2015; @wiessner_collective_2019; @pupu_challenges_2018; @hagen_leadership_2019]. -->

@@

<!-- SUMMARY AND SUGGEST LEADERSHIP BE PART OF OTHER GROUP RESEARCH -->
<!-- These exploratory results provide an empirical foundation for future research on leadership across or within diverse cultural settings and levels of social organization. Political leadership is common across a broad range of populations, including among relatively small-scale and egalitarian hunter-gatherers. Among hunter-gatherers, however, leaders more frequently emerge within residential subgroups and kin groups. Given the vast majority of human evolutionary history involved subsistence economies characterized by hunting and gathering and greater residential mobility [@lee_cambridge_1999; @marlowe_hunter-gatherers_2005], these results suggest leadership within kin groups and local subgroups likely has a long evolutionary history. Despite a universal male-bias in political leadership at the community level, women are expected to have greater decision-making responsibilities and adopt leadership roles within families and local groups [@brown_her_1985; @garfield_evolutionary_2019; @hagen_leadership_2019; @smith_obstacles_2020; @von_rueden_sex_2018].  -->

@@

**WEIRD leadership costs and benefits**. Conceptualizing leadership as a possible solution to problems associated with collective goods and actions [e.g., @olson_logic_1965; @glowacki_leadership_2015; @hooper_theory_2010; @pietraszewski_evolution_2020], where group members pool resources and share costs of public goods to receive a net benefit from a cooperative endeavor, leaders incur costs associated with managing contributions, coordinating behaviors, and broadcasting information to maintain coordination and cooperation. In this framework, leaders and followers voluntary participate in a cooperative enterprise with the expectation that benefits will outweigh the associated costs. @rich_dynamics_1980 adopts this framework of collective costs and benefits to understand leadership and group dynamics in WEIRD urban neighborhood organizations, which can be viewed as self-interested groups that face group-wide problems and necessarily share public benefits. The costs associated with leadership in neighborhood organizations follow those associated with other collective enterprises (e.g., investments of time and energy in management and maintenance of group coordination). @rich_dynamics_1980 conceptualizes four categories of the benefits associated with neighborhood organization. Leaders of neighborhood organizations (1) may receive benefits of the public goods produces from collective actions (as other non-leaders would as well); (2), may receive selective benefits from surpluses produced through collective actions; (3) may receive incidental benefits from their leadership role; and (4) leaders of neighborhood organizations may enjoy their service in the position. @rich_dynamics_1980 interviewed 54 officers from 11 metropolitan neighborhood organizations in the Midwest US. Results from extensive interviews suggest personal fulfillment and enjoyment are the main motivational drivers of participation in leadership roles of neighborhood organizations, suggesting the process of leadership for individuals is distinct from theoretical models rooted in economics. Therefore, a primary benefit of leadership, in some contexts, is likely to be intrinsic personal fulfillment and the opportunity to execute personal values of community service.
